# Reproducible Research: Peer Assessment 1

## Loading library
```{r}
library(data.table)
library(fasttime)
library(ggplot2)
library(gridExtra)
```

## Getting data
```{r}
if (!file.exists('activity.csv')) unzip("activity.zip")
```

## Loading and preprocessing the data
```{r}
testData <- fread("activity.csv",nrows=1000)
summary(testData)
head(testData)
testData$date <- fasttime::fastPOSIXct(testData$date)
summary(testData$date)

data <- fread("activity.csv")
data$date <- fasttime::fastPOSIXct(data$date)
summary(data)
orgData <- data
data <- na.omit(data) #Ommited missing value
```


## What is mean total number of steps taken per day?
### A histogram of the total number of steps taken each day
```{r}
obsGroupByPerDay = orgData[,list(total_steps=sum(steps)),by=date] #NEED: V1 column name -> MeanOfSteps
p <- ggplot(data=obsGroupByPerDay) + geom_histogram(aes(x=total_steps,fill=..count..)) +
  xlab("Steps")
print(p)
```

### The mean total number of steps taken per day
```{r}
meanStepsPerDay = orgData[,mean(steps),by=date]
meanStepsPerDay
```

### The median total number of steps taken per day
```{r}
medianStepsPerDay = data[,median(steps,na.rm=T),by=date]
medianStepsPerDay
```

## What is the average daily activity pattern?
### A time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
meanStepByInterval = data[,mean(steps),by=interval]
ggplot(data=meanStepByInterval) + geom_line(aes(x=interval,y=V1)) + ylab("mean of steps")
```

### 5-minute interval contains the maximum number of steps
```{r}
meanStepByInterval$interval[which.max(meanStepByInterval$V1)]
835/60 # At 2pm
```

## Imputing missing values
### the total number of missing values
#### Per day
```{r,missingValuesPerDay}
orgData[,list(count=sum(is.na(steps))),by=date][count>0]
```
#### All day
```{r,missingValuesAllDay}
summary(orgData) # 2304
```

### Filling in all of the missing values in the dataset
#### Filling missing intervals in a day
*Strategy: Use steps' mean of a day to fill missing values of step that day*
```{r}
fillMissingIntervalInADayData = orgData
naPostition = which(is.na(fillMissingIntervalInADayData$steps))
naFilled = merge(fillMissingIntervalInADayData[naPostition,list(date)],medianStepsPerDay,by="date",all.x=TRUE) #Merge with median per day
fillMissingIntervalInADayData$steps[naPostition] = naFilled$V1
summary(fillMissingIntervalInADayData) # 2304
```

#### Filling a day which have missing intervals
*Strategy: Use mean of steps group by intervals to fill missing values of steps belong to the intervals*
```{r}
#Duplicate data
fillADayMissingInvervals = fillMissingIntervalInADayData
#Find NA row position
naPostition = which(is.na(fillADayMissingInvervals$steps))
#Filled NA by interval
naFilledValues = merge(fillADayMissingInvervals[naPostition,list(interval)],meanStepByInterval,by="interval",all.x=TRUE)
#Assign back
fillADayMissingInvervals$steps[naPostition] = naFilledValues$V1
summary(fillADayMissingInvervals)
```
### A new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
filledData = fillADayMissingInvervals
```

### Compare before and after filling missing values
```{r,compareBetweenBeforeAndAfterFilling}
totalStepEachDay = filledData[,list(total_steps=sum(steps)),by=date]
before = ggplot(data=obsGroupByPerDay) + 
  geom_histogram(aes(x=total_steps,fill=..count..),binwidth=500) + 
  ggtitle("Before filling")
after = ggplot(data=totalStepEachDay) + 
  geom_histogram(aes(total_steps,fill=..count..),binwidth=500) + 
  ggtitle("After filling")
grid.arrange(before,after)
```

*Don't see much difference here*

## Are there differences in activity patterns between weekdays and weekends?
1. Create dayTypes variable indicate day as weekends or weekdays
```{r}
filledData$dayTypes = as.factor(filledData[,grepl("Sunday|Staturday",weekdays(date))])
levels(filledData$dayTypes) <- c("weekdays","weekends")

```
2. Make panel plot show number of steps for each interval split by dayTypes
```{r}
filledDataSumStepByDayTypesAndInterval = filledData[,list(sum=sum(steps)),by=list(dayTypes,interval)]
ggplot(data=filledDataSumStepByDayTypesAndInterval) +
  geom_line(aes(x=interval,y=sum)) +
  facet_grid(dayTypes ~ . ) +
  theme(strip.text.y = element_text(size = 20)) +
  ylab("Number of steps")
```

*Yes much difference, seem lazier at weekends*
